Effect of Earnings and Dividends on Stock Prices

### 2024-04-12 outputs/paper/paper.qmd
intro and abstract

---
title: "Poopy But hole"
subtitle: "Analysis of the Relationship Between Stock Price, Earnings Per Share, and Dividends of the Largest 50 North American Companies from 2013 to 2023"
author: Diana Liu
thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Relationship-Between-Stock-Prices-and-Common-Financial-Metrics"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction

# Data
Public companies whose stock are traded on exchanges, must disclose certain financial metrics in their quarterly and yearly financial statements. Investors then use these metrics to help inform their decision when selecting companies to invest in. Some of the most important metrics investors look at are Earnings Per Share, Net Income, and Dividends [@CiteShakespeare]. Investments take the form of buying shares or stocks which allows the investor to own a fraction of the company.

The data used for analysis is obtained through Walton Research Data Services (WRDS) by the University of Pennsylvania. WRDS provides access to Compustat, a database of financial, statistical, and market information on global companies since 1962. Our data comes from Compustat's North America Fundamentals Annual database, which contains both financial data from all public North American companies, collected by Compustat from each company's annual financial statements or from stock exchanges (@fig-Analysis-Data-Sample). From this data set, we extract the price, EPS, net income, dividends, TIC, and Year of the top 50 largest companies by market capitalization, a measure of the size of the company in terms of total value, for the past 10 years (2013-2023). This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, @CiteKnitr, @Citegt, and @CitegrDevices.

@@

# Conclusion
We have determined that there is a positive linear relationship between EPS and stock price, and a negative relationship between dividends and stock price which reduces the affect of EPS on price. These relationships can be used by investors to predict stock price of a given company based on EPS and dividends. The investor can then earn money by trading stocks accordingly, buying when the price is predicted to increase and selling when it will fall. Managers of companies can also use our results to make more informed decisions, balancing the objectives of pleasing current shareholders, investing excess cash into promising projects, and maintainging the attractiveness of their stock for new investors.

>>>>

---
title: "Effect of Earnings and Dividends on Stock Prices"
subtitle: "Analysis of the Relationship Between Stock Price, Earnings Per Share, and Dividends of the Largest 50 North American Companies from 2013 to 2023"
author: Diana Liu
thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Relationship-Between-Stock-Prices-and-Common-Financial-Metrics"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction
Public companies whose stock are traded on exchanges, must disclose certain financial metrics in their quarterly and yearly financial statements. Investors can then use these metrics to help inform their decision when selecting companies to invest in, their investment of buying shares represent a fractional ownership of the company. Some of the most important metrics investors look at are Net Income, Earnings Per Share (EPS), and Dividends [@CiteShakespeare]. In the data section, we discuss how EPS is a general measure of a firm's profitability, dividends are cash distributed to investors, and stock prices are determined by both the intrinsic value of the firm as well as market pressures like demand for the stock. Our estimand is the relationship between EPS and stock price as well has how this relationship is effected by dividends.

In the model section, we construct two models with model 1 being a simple linear model between EPS and stock price to determine their relationship, and model 2 being a multivariable model that groups companies by those who paid dividends and those who did not. From the results section, model 1 found that there is a positive linear relationship between EPS and stock price with every dollar increase in EPS correlating with a 26.59 USD increase in stock price. We also determined that there is a negative relationship between paying dividends and stock price. In Model 2, if the company paid dividends, every dollar increase in EPS correlates with only a 25.59 USD increase in stock price.

We have determined that there is a positive linear relationship between EPS and stock price, and a negative relationship between dividends and stock price which reduces the affect of EPS on price. These relationships can be used by investors to predict stock price of a given company based on EPS and dividends. The investor can then earn money by trading stocks accordingly, buying when the price is predicted to increase and selling when it will fall. Managers of companies can also use our results to make more informed decisions, balancing the objectives of pleasing current shareholders, investing excess income into promising projects, and maintaining the attractiveness of their stock for new investors. Further analysis should be conducted using a more diverse data set in terms of geography and company size in order to confirm these relationships for all companies. Another weakness is that stock price is also influenced by categorical factors like a company's strategic direction and competitive advantage which are variables that could not be included in our analysis. Future efforts should attempt to determine the effects of these categorical variables on stock price and how it interacts with that of EPS and dividends.

# Data
The data used for analysis is obtained through Walton Research Data Services (WRDS) by the University of Pennsylvania. WRDS provides access to Compustat, a database of financial, statistical, and market information on global companies since 1962. Our data comes from Compustat's North America Fundamentals Annual database, which contains both financial data from all public North American companies, collected by Compustat from each company's annual financial statements or from stock exchanges (@fig-Analysis-Data-Sample). From this data set, we extract the price, EPS, net income, dividends, TIC, and Year of the top 50 largest companies by market capitalization, a measure of the size of the company in terms of total value, for the past 10 years (2013-2023). This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, @CiteKnitr, @Citegt, and @CitegrDevices.

@@

# Conclusion
